<!-- R Commander Markdown Template -->

Homework 3 Regression
=======================

### Abby Mutic

### 02/15/17
 
```{r}
Dataset <- 
  
  read.table("C:/Users/AHENDE8/Documents/R class- big data/Regression-1/White_wines.csv",
  
   header=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)
```


```{r}
library(abind, pos=15)
```


```{r}
library(e1071, pos=16)
```


```{r}
numSummary(Dataset[,c("alcohol", "chlorides", "citric.acid", "density", 
  "fixed.acidity", "free.sulfur.dioxide", "pH", "quality", "residual.sugar", 
  "sulphates", "total.sulfur.dioxide", "volatile.acidity")], statistics=c("mean", 
  "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,.75,1))
```
#Based on summary statistics, there are 4898 observations with no missing data reported on any of the variables. There are 12 variables total.
#Total sulfur dioxide and free sulfur dioxide have large SD and ranges possibly indicating some outliers could exist.

#Now look at univariate distributions with histograms.

```{r}
with(Dataset, Hist(alcohol, scale="frequency", breaks="Sturges", col="darkgray"))
```

```{r}
with(Dataset, Hist(chlorides, scale="frequency", breaks="Sturges", col="darkgray"))
```

```{r}
with(Dataset, Hist(citric.acid, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```

```{r}
with(Dataset, Hist(density, scale="frequency", breaks="Sturges", col="darkgray"))
```

```{r}
with(Dataset, Hist(fixed.acidity, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```

```{r}
with(Dataset, Hist(free.sulfur.dioxide, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```

```{r}
with(Dataset, Hist(pH, scale="frequency", breaks="Sturges", col="darkgray"))
```

```{r}
with(Dataset, Hist(quality, scale="frequency", breaks="Sturges", col="darkgray"))
```

```{r}
with(Dataset, Hist(residual.sugar, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```

```{r}
with(Dataset, Hist(sulphates, scale="frequency", breaks="Sturges", col="darkgray"))
```

```{r}
with(Dataset, Hist(total.sulfur.dioxide, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```

```{r}
with(Dataset, Hist(volatile.acidity, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
#About half of the univariate histograms look right skewed. The obvious normal distributions are for the variables: alcohol, density, fixed acidity, pH. sulfates, total sulfur dioxide.  

#Now look at multivariate graphs with scatterplot matrix.


```{r}
scatterplotMatrix(~alcohol+chlorides+citric.acid, reg.line=FALSE, smooth=FALSE, 
  spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), id.n=0, diagonal = 
  'density', data=Dataset)
```

```{r}
scatterplotMatrix(~density+fixed.acidity+free.sulfur.dioxide, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), id.n=0, 
  diagonal = 'density', data=Dataset)
```

```{r}
scatterplotMatrix(~pH+quality+residual.sugar, reg.line=FALSE, smooth=FALSE, 
  spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), id.n=0, diagonal = 
  'density', data=Dataset)
```

```{r}
scatterplotMatrix(~sulphates+total.sulfur.dioxide+volatile.acidity, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), id.n=0, 
  diagonal = 'density', data=Dataset)
```
#look at a few bivariate scatterplots

```{r}
scatterplot(quality~alcohol, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~chlorides, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~citric.acid, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~density, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~fixed.acidity, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~free.sulfur.dioxide, reg.line=FALSE, smooth=FALSE, spread=FALSE,
   boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~pH, reg.line=FALSE, smooth=FALSE, spread=FALSE, boxplots=FALSE, 
  span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~residual.sugar, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~sulphates, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

```{r}
scatterplot(quality~total.sulfur.dioxide, reg.line=FALSE, smooth=FALSE, 
  spread=FALSE, boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  data=Dataset)
```

```{r}
scatterplot(quality~volatile.acidity, reg.line=FALSE, smooth=FALSE, spread=FALSE, 
  boxplots=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), data=Dataset)
```

#Build linear regression model with all variables in the model

```{r}
Regmodel1 <- 
  lm(quality~alcohol+chlorides+citric.acid+density+fixed.acidity+free.sulfur.dioxide+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=Dataset)
summary(Regmodel1)
```

#Drop all nonsiginificant variables.
```{r}
Regmodel2 <- lm(quality ~ alcohol+density+fixed.acidity+ free.sulfur.dioxide+pH+residual.sugar, data=Dataset)
summary(Regmodel2)
```
# All variables are significant. Adjusted R squared is 23%. This means ~23% of y (quality) is due to the other variables in the model.

```{r}
stargazer(Regmodel1, Regmodel2, title = "Comparison of 2 regression outputs", type = "html")
```

#Look at output in text form
```{r}
stargazer(Regmodel1, Regmodel2,
          title = "Comparison of 2 Regression outputs",
          type = "text")
```

```{r}
residualPlots(Regmodel2)
```
#Variables alcohol and pH are not near zero so I will remove them #from the model.

```{r}
Regmodel3 <- lm(quality ~ density+fixed.acidity+ free.sulfur.dioxide+residual.sugar, data=Dataset)
summary(Regmodel3)
```
#Adjusted R squared lowered to 19%. Now will look at residual plots
```{r}
residualPlots(Regmodel3)
```
#All the test statistics are now significant.
#Now run the rest of the diagnostic statistics.

```{r}
avPlots(Regmodel3, id.n=3, id.cex=0.7)
```
```{r}
qqPlot(Regmodel3, id.n=3)
```

```{r}
influenceIndexPlot(Regmodel3, id.n=3)
```
```{r}
outlierTest(Regmodel2)
```

```{r}
ncvTest(Regmodel3)
```

```{r}
vif(Regmodel3)
```
 The final model includes density, fixed acidity, free sulfur dioxide, and residual sugar to predict the best qualities of white wine and/or preferences of white wine. The implications of this statistical analysis can inform wine and grape producers about the preferences of their consumers and hopefully increase sales for the wine companies.
 This Github repository is located at 
